---
title: "Synthetic Benchmark I"
author: "Gary Glaze"
date: "Friday, January 02, 2015"
output: pdf_document
---
The proposed synthetic benchmark combines Vanguard Value Index Admiral Fund (VVIAX) 100% long with a short position in the Vanguard Growth Index Admiral Fund (VIGAX). Target portfolio return equals Rvviax + 0.75*(Rvviax - Rvigax).

Vanguard created both component funds in November of 2000. This document examines their monthly returns  from January 2001 to December 2014. It also compares their performance to the S&P 500.

```{r,echo=FALSE,results='hide',message=FALSE}
options(digits=4, width=70)

library(PerformanceAnalytics) # for return and risk analysis
library(zoo) # Zeilie's ordered observations
library(tseries) # time series functions

# get the adjusted closing prices from Yahoo!
GSPC.prices = get.hist.quote(instrument="^gspc", start="2001-01-31",
                              end="2014-12-31", quote="AdjClose",
                              provider="yahoo", origin="1970-01-01",
                              compression="m", retclass="zoo")
# change class of time index to yearmon which is appropriate for monthly data
# index() and as.yearmon() are functions in the zoo package 
index(GSPC.prices) = as.yearmon(index(GSPC.prices))

# Vanguard Value Index Admiral Fund (inception date: 11/13/2000)
VVIAX.prices = get.hist.quote(instrument="vviax", start="2001-01-31",
                            end="2014-12-31", quote="AdjClose",
                             provider="yahoo", origin="1970-01-01",
                             compression="m", retclass="zoo")
index(VVIAX.prices) = as.yearmon(index(VVIAX.prices))
                             
# Vanguard Growth Index Admiral Fund (inception date: 11/13/2000)
VIGAX.prices = get.hist.quote(instrument="vigax", start="2001-01-31",
                             end="2014-12-31", quote="AdjClose",
                             provider="yahoo", origin="1970-01-01",
                             compression="m", retclass="zoo")
index(VIGAX.prices) = as.yearmon(index(VIGAX.prices))

# create merged price data
lab5Prices.z = merge(GSPC.prices, VVIAX.prices, VIGAX.prices)
# rename columns
colnames(lab5Prices.z) = c("S&P500", "VVIAX", "VIGAX")

# calculate cc returns as difference in log prices
lab5Returns.z = diff(log(lab5Prices.z))
```

**Initial Monthly Returns**
```{r,echo=FALSE}
head(lab5Returns.z)

```

**Plots of Monthly Returns**

```{r,echo=FALSE}
# 3-panel plot (each y axis has different scale)
plot(lab5Returns.z,col="blue", lwd=2, main="Monthly CC Returns on 3 Assets")
# 
# CC returns all on same graph (function chart.Bar from PerformanceAnalytics)
chart.Bar(lab5Returns.z, legend.loc="bottom", main="Monthly CC Returns on 3 Assets", ylab="Percent")
```

\pagebreak

**Plot of Cumulative Returns**

```{r,echo=FALSE}
# cumulative return plot - must use simple returns and not cc returns for this
# use PerformanceAnalytics function chart.CumReturns()
chart.CumReturns(diff(lab5Prices.z)/lag(lab5Prices.z, k=-1), 
                 legend.loc="topleft", wealth.index=TRUE,
                 main="Future Value of $1 Invested")
```

```{r,echo=F,message=FALSE,results='hide'}
# Create matrix of return data. Some core R functions don't work correctly with zoo objects. 
ret.mat = coredata(lab5Returns.z)
```

\pagebreak

**All funds show significant negative skew. Distributions are approximately normal for values greater than negative one standard devation.**

*Vanguard Value Fund has greatest negative skew*

```{r,echo=FALSE}
# here are the 4 panel plots
par(mfrow=c(2,2))
    hist(ret.mat[,"VVIAX"],main="VVIAX monthly returns",
	     xlab="VVIAX", probability=T, col="slateblue1")
	boxplot(ret.mat[,"VVIAX"],outchar=T, main="Boxplot", col="slateblue1")
	plot(density(ret.mat[,"VVIAX"]),type="l", main="Smoothed density",
           xlab="monthly return", ylab="density estimate", col="slateblue1")
	qqnorm(ret.mat[,"VVIAX"], col="slateblue1")
	qqline(ret.mat[,"VVIAX"])
par(mfrow=c(1,1))
```

\pagebreak

*Vanguard Growth Fund*

```{r,echo=FALSE}
par(mfrow=c(2,2))
	hist(ret.mat[,"VIGAX"],main="VIGAX monthly returns",
	     xlab="VIGAX", probability=T, col="slateblue1")
	boxplot(ret.mat[,"VIGAX"],outchar=T, main="Boxplot", col="slateblue1")
	plot(density(ret.mat[,"VIGAX"]),type="l", main="Smoothed density",
	     xlab="monthly return", ylab="density estimate", col="slateblue1")
	qqnorm(ret.mat[,"VIGAX"], col="slateblue1")
	qqline(ret.mat[,"VIGAX"])
par(mfrow=c(1,1))
```

\pagebreak

*S&P 500 shows less negative skew*

```{r,echo=FALSE}
par(mfrow=c(2,2))
	hist(ret.mat[,"S&P500"],main="S&P500 monthly returns",
	     xlab="S&P500", probability=T, col="slateblue1")
	boxplot(ret.mat[,"S&P500"],outchar=T, main="Boxplot", col="slateblue1")
	plot(density(ret.mat[,"S&P500"]),type="l", main="Smoothed density",
	     xlab="monthly return", ylab="density estimate", col="slateblue1")
	qqnorm(ret.mat[,"S&P500"], col="slateblue1")
	qqline(ret.mat[,"S&P500"])
par(mfrow=c(1,1))
```

**Summary Statistics**

Note that all distributions have significant negative skewness and positive kurtosis. Negative skew means simply that the data contain more extreme negative values than a normal distribution. Positive Kurtosis indicates that returns group less tightly around the mean than for a normal distribution.

```{r}
table.Stats(lab5Returns.z)
```

\pagebreak

**Correlation Plots**

All funds are highly correlated. The Vanguard Growth and Vanguard Value funds are somewhat less correlated with each other than with the S&P 500.

```{r,echo=FALSE}
pairs(ret.mat, col="slateblue1", pch=16)
```

**Correlation Matrix**

```{r,echo=FALSE}
cor(ret.mat)
```

\pagebreak

**Autocorrelation**

Plots show that monthly returns are not serially autocorrelated.

```{r,echo=FALSE}
par(mfrow=c(3,1))
    acf.vviax = acf(ret.mat[,"VVIAX"], main="VVIAX")
	acf.vigax = acf(ret.mat[,"VIGAX"], main="VIGAX")
	acf.sp500 = acf(ret.mat[,"S&P500"], main="S&P500")
par(mfrow=c(1,1))
```
